---
title: "MATH 3070 Lab Project 2"
author: "Vishwa Murugappan"
date: "September 1, 2022"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 1.7)

*The `rivers` (**UsingR**) or any other data set is available after loading the package **UsingR**. Load the package, and inspect the data set. Scan the values to find the largest one.*

The dataset is not available by default in R. The UsingR package must be imported to access it. The max() function can retrieve the largest value in the dataset.

```{r}
# Requires that the UsingR pakcage be installed. This is necessary to get the rivers data.
require(UsingR)

# Import data
data(rivers)

# Print 6 key values that descirbe the dataset and its spread
cat("Dataset Inspection and Summary:","\n")
summary(rivers)

# Print maximum value in the dataset
cat("\n", "Maximum is:", max(rivers))
```

## Problem 2 (Verzani problem 2.4)

*Create the following sequences, using `:`, `seq()`, or `rep()` as appropriate:*

1.  `"a" "a" "a" "a" "a" "a"`

We want 6 a's in a row.

```{r}
# rep() allows you to repeat a string a specified amount of times
rep("a",6)
```

2.  `1 3 ... 99` *(the odd numbers)*

We want a sequence beginning with 1, ending with 99, and increasing in increments of 2.

```{r}
# seq() allows for a a sequence to be generated from one number to another one with a specified
# amount of spacing
seq(1, 99, 2)
```

3.  `1 1 1 2 2 2 3 3 3`

We want a sequence going from 1 to 3 with each number being repeated 3 times.

```{r}
# rep() also has an optional "each" parameter that lets you pick a number of times you want a
# value to show up before mvoing on to the next one
rep(1:3, each = 3)
```

## Problem 3 (Verzani problem 2.3)

*Let our small data set be* `2 5 4 10 8`.

1.  *Enter this data into a data vector `x`.*

Vectors are the result of a combinations of values, like the dataset presented earlier.

```{r}
# c() allows for values to be combined into a vector or list
# We can make the dataset using it
x <- c(2,5,4,10,8)
```

2.  *Find the square of each number.*

We can raise the dataset to the power of 2.

```{r}
# When an operation is applied on a list, the operation is applied to all of its component
# values
x ^ 2
```

3.  *Subtract 6 from each number.*

We can subtract 6 from the dataset.

```{r}
# Same as before
# When an operation is applied on a list, the operation is applied to all of its component
# values
x - 6
```

4.  *Subtract 9 from each number and then square the answer.*

We can subtract 9 from the dataset and then raise it to the power of 2.

```{r}
# Same as before; slightly more complicated operation
# When an operation is applied on a list, the operation is applied to all of its component
# values
(x - 9) ^ 2
```
