---
title: "MATH 3070 Lab Project 8"
author: "Your Name"
date: "October 18, 2017"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

## Problem 1 (Verzani problem 3.2)

*For the `michelson` (**MASS**) data set, produce a density plot comparing `Speed` between Experiments 1 and 2.*

```{r, tidy=TRUE, error=TRUE}
# Importing required libraries
library(MASS)
```

```{r, tidy=TRUE, error=TRUE}
# Using subset() to fulter dataset by experiment value (1 and 2)
expt1 = subset(michelson, Expt == 1)
expt2 = subset(michelson, Expt == 2)

# plot(density()) will plot the density plots of speeds for each experiment
plot(density(expt2$Speed), lty = 1, main = "Density Plots of Speed for Experiments 1 and 2")
lines(density(expt1$Speed), lty = 2)
# Legend necessary to distinguish between which line represents which experiment
legend(930, 0.0055, legend=c("Expt 1 Speed Density", "Expt 2 Speed Density"),, lty=2:1, cex=0.7)
```

This plot shows the distribution of the values of speed for experiments 1 and 2 in the michelson dataset.

## Problem 2 (Verzani problem 3.4)

*Three students record the time spent on homework per class. Their data is:*

| Student |   1 |   2 |   3 |   4 |   5 |
|:--------|----:|----:|----:|----:|----:|
| Marsha  |  25 |   0 |  45 |  90 |   0 |
| Bill    |  30 |  30 |  30 |  30 |     |
| Holly   |  15 |   0 |  90 |   0 |     |

*Use a list to store these values. Then create a boxplot to compare. (You must use `boxplot()`'s formula interface for this problem.)*

```{r, tidy=TRUE, error=TRUE}
# This list contains a vector of completion times and the students with the number of completions "belonging" to them. This was created using the rep() function with the times parameter
hw_time = list(time = c(25, 0, 45, 90, 0, 30, 30, 30, 30, 15, 0, 90, 0),
          students = rep(c("Marsha", "Bill", "Holly"),
                       times = c(5,4,4)))
# completion_time ~ student made use of the formula interface in boxplot() because it shows a response ~ predictor relationship between the two
boxplot(time ~ students, data = hw_time)
```

Bill has no variation in the amount of time it took to complete homework for each of his classes, so the median, 1st quartile, and 3rd quartile are all 30 for him.

Holly has the lowest median out of the three (7.5). However, she has an outlier value of 90, which causes her to have a high 3rd quartile value (52.5). Because she spends no time on some assignments, she has a 1st quartile value of (0). Overall, she has the largest IQR.

Marsha has a median value of 25. She also has an outlier value of 90, which causes her to have a high 3rd quartile value (67.5). Because she spends no time on some assignments, she has a 1st quartile value of (0).

## Problem 3 (Verzani problem 3.5)

*A group of nursing students take turns measuring some basic assessments. Their data is:*

|          | Temp | Pulse | Systolic | Diastolic |
|----------|------|-------|----------|-----------|
| Jackie   | 98.2 | 96    | 134      | 90        |
| Florence | 98.6 | 56    | 120      | 80        |
| Mildred  | 98.2 | 76    | 150      | 95        |

*Create a data frame of these values. Will `plot()` and `boxplot()` produce the same graphic?*

```{r, tidy=TRUE, error=TRUE}
# This creates the dataframe with each variable and its values.
nursing = data.frame (Temp  = c(98.2,98.6,98.2),
                  Pulse = c(96,56,76),
                  Systolic = c(134,120,150),
                  Diastolic = c(90,80,95))
# This is not necessary to complete the problem, but to replicate the data provided above, the observations were named using rownames() as well.
rownames(nursing) <- c("Jackie", "Florence", "Mildred")

# Plots provided are not the same
plot(nursing)
boxplot(nursing)
```

The plot() function shows pair plots of each variable (axes change on individual plots). The boxplot() function shows box plots of each variable, all with the same vertical axis scale.

## Problem 4 (Verzani problem 3.8)

*The second argument to `split` can be a list of factors. The result is that all interactions (possible combinations) are used for the groups. In the `ToothGrowth` data set, growth (`len`) is measured for two types of supplements (`supp`) and three doses (`dose`). Split this `len` value into 6 groups.*

```{r, tidy=TRUE, error=TRUE}
# This can be done by passing in a list with the supp and dose variables as the f (factor) parameter.
split(ToothGrowth$len, f = list(ToothGrowth$supp, ToothGrowth$dose))
```

There are six groups above with the different possible combinations of supp values (OJ and VC) as well as dose values (0.5, 1, and 2).
