---
title: "MATH 3070 Lab Project 1"
author: "Vishwa Murugappan"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

*Remember: I expect to see commentary either in the text, in the code with comments created using `#`, or (preferably) both! **Failing to do so may result in lost points!***

*Note: The dollar signs, back slashes, frac, etc. are all part of the typesetting language* $\LaTeX$. They are used to create good-looking equations upon the knit. However, they are not R code. Thus the goal of this problem is to essentially translate from $\LaTeX$ to R. `frac` will need to be replaced with the usual divide sign, for example

## Problem 1 (Verzani problem 1.1)

*Use R as you would a calculator to find numeric answers to the following:*

1.  $1 + 2(3 + 4)$

***I used R as a calculator to solve the math problem below.***

```{r}
# This problem requires the usage of the multiplication and addition operators.
1 + 2*(3 + 4)
```

2.  $4^3 + 3^{2 + 1}$

***I used R as a calculator to solve the math problem below.***

```{r}
# This problem requires the usage of the multiplication, addition, and caret (exponent) operators.
4^3 + 3^(2 + 1)
```

## Problem 2 (Verzani problem 1.4)

*Use R to compute the following:*

$$\frac{0.25 - 0.2}{\sqrt{0.2(1-0.2)/100}}$$

***This Problem is more difficult than the previous ones because it requires you to figure out how to represent square roots in R code.***

```{r}
# I used R as a calculator to solve the math problem above.
# I noticed that sqrt() is the function that is used in R to take a square root.
(0.25 - 0.2)/sqrt((0.2*(1-0.2)/100))
```

## Problem 3 (Verzani problem 1.5)

*Assign the numbers 2 through 5 to different variables, then use the variables to multiply all the values.*

***Defined variables in R can be multiplied the same way as numbers.***

```{r}
# I used '<-' to assign values to all the variables.
# Then, I used the multiplication operator to multiple all of them together.
two <- 2
three <- 3
four <- 4
five <- 5
two * three * four * five
```
